\section{Requirements}

\subsection{Software Requirements}

\subsubsection*{Software requirements for this project include:}


\begin{itemize}
    \item Python 3.10
    \item Pyboy
    \item Stablebaselines3
    \item Tensorboard
    \item Wandb
    \item Git
    \item LaTeX
    \item Pokémon Red Rom
\end{itemize}

\subsubsection*{Python}
Python is the main programming language used for this project. Python is a high-level, general-purpose programming language that is widely used in the field of machine learning. The implementation of the environment using the gymnasium package was written in Python, where the agents were created using the Stablebaselines3 implementation of A2C, DQN, and PPO.

\subsubsection*{Pyboy}
Pyboy is a Gameboy emulator written in Python. Pyboy is used to run the Pokémon Red ROM to interact with the game, where each button in the Gameboy emulator was mapped to a key on the keyboard. Pyboy is used to create the environment by wrapping it around a gymnasium environment, where the information on the state of the environment is taken from Pyboy via RAM readings or information on the screen. Pyboy was chosen as the Gameboy emulator within Python as it was the most well-documented Gameboy emulator.

\subsubsection*{Stablebaselines3}
Stablebaselines3 is a library of RL algorithms implemented in Python. Stablebaselines3 is used to create agents that interact with the environment and was specifically chosen because it was a very straightforward-to-use library that worked well with the gymnasium environment library.

\subsubsection*{Tensorboard}
Tensorboard is a tool used to visualize the training of the agents through the use of graphing measurable metrics. It was used to compare the performance of the trained models. Tensorboard was not necessary for this project but it was a very useful tool to visualize the progression of the training of the agents. Tensorboard was also a recommended tool that was supported by both stablebaselines3 algorithms and gymnasium environments.

\subsubsection*{Wandb}
Wandb is another tool used to visualize the training of the agents similar to tensorboard. Wandb can do everything tensorboard does with the added benefit of being able to save console logs, code used to run the script, and track hardware performance during training. It was used together with tensorboard for the purpose of documenting my research.

\subsubsection*{Git}
Git is a version control system used to track changes and to keep a backup of the code on GitHub where there is a link to the project repository. It served no core purpose to this project but is a programming standard. 

\subsubsection*{LaTeX}
LaTeX is a typesetting system that is widely used for the publication of documents. LaTeX is used to create the final report of this research project.

\subsubsection*{Pokémon Red Rom}
The Pokémon Red Rom is the game that is loaded into Pyboy which makes the environment the agent interacts with. The Rom is not included in the repository and must be downloaded separately. A legal copy of the game is owned by the author of this project prior to the start of this research.


\subsubsection*{Python Package requirements for this project include: }

\begin{multicols}{2}
\begin{itemize}
    \item uuid
    \item json
    \item pathlib
    \item numpy
    \item skimage
    \item matplotlib
    \item pyboy
    \item mediapy
    \item gymnasium
    \item asyncio
    \item websockets
    \item os
    \item datetime
    \item stable-baselines3
    \item tensorboard
    \item wandb
    \item einops
\end{itemize}
\end{multicols}

\subsection{Project Recreateability}\label{subsec:Hardware}

Hardware used for this project includes:
\begin{itemize}
    \item Linux: Arch Kde Plasma 5.23.3, 64 GB RAM, AMD Ryzen 7 7700X, Nvidia RTX 4070 ti
    \item 64 GB RAM minimum to run 11 parallel agents training on the environment per episode

\end{itemize}

\subsection{Recreateability of Research}

Public GitHub repository for the project: \url{https://github.com/liamo3323/FYP_PokémonRed_RL}

Experiments mentioned in this research can be prepared by pulling the GitHub repo mentioned and placing the ``PokemonRed.gb'' file in the root directory. The ``PokemonRed.gb'' file was not included in the repository due to legal reasons. 

The agent can be trained running the ``run\_baseline'' file in the root directory, where the settings for each algorithm are specified in the file name. The option to train an agent from scratch or to load a pre-trained model is available. The trained model will then be saved within the ``sessions'' directory.

Wandb visualization is enabled within the ``run\_baseline.py'' file and tensorboard metrics will saved within the model's directory. To visualize the performance of the agent, the wandb project name will need to be updated.

Pre-trained models for 40 million timesteps are saved within the ``40mil\_Saved\_Sessions'' directory. The pre-trained models can be loaded and pre-viewed using ``run\_pretrained\_v2.py''. Further instructions can be read in the README.md file in the root directory or GitHub repository.