\section{Ethical Consideration}

When conducting research in any field, taking the ethical and legal considerations of the work conducted is integral to the research conducted. Throughout the work conducted in this project, all legal and ethical considerations have been taken into account. This was ensured through the guidance of my supervisor and by following the BCS Code of Conduct. In addition, following the ethical guidance of the University of Surrey, the Self-Assessment for Governance and Ethics (SAGE) was completed to ensure that the research conducted was ethical and legal.


\subsection{Public Interest - Do no harm}

Throughout this project, I have respected the legality, privacy and confidentiality of the data used. The data used was the Pokémon Red ROM, which was used to create the environment for the reinforcement learning agent. However, there are potential legal issues regarding the use of the ROM, as it is owned by Nintendo. However, the use of the ROM was for educational purposes and not for commercial gain. Moreover, a physical copy of the video game is owned by the researcher, which implies that the use of the ROM is not illegal.

\subsection{Informed Consent}

The research conducted in this project did not include any human participation or data. Therefore, there was no need for informed consent to be obtained.

\subsection{Confidentiality of Data}

No private, personal or confidential data was used for the purpose of this research. The data that was used was the Pokémon Red ROM, which was used to create the environment for the reinforcement learning agent. The data was not used for any other purpose than to create the environment for the agent to learn from. Therefore, the only type of data used for this research was directly used for the environments and no personal data was used for this research.

The limited data that is used is stored within the environment and only processed for the purpose of training the agent. The data in question is also not directly remembered or stored within the agent, only an association of how to react to the data within the environment is stored. Furthermore, no data is stored outside of the environment. 

\subsection{Social Responsibility}

The purpose of this research holds no real-world implications and the development of the reinforcement learning agent is limited to the environment of Pokémon Red, which is a game and holds no real-world implications. The game is also marketed towards children, which implies that it is not harmful to any individual or group of individuals. Moreover, the greater purpose of this research is to assist the field of multi-objective reinforcement learning by providing more data for benchmarking and comparison, which holds just as many potential positives as negative implications.

\subsection{British Computer Society Code of Conduct}

The British Computer Society (BCS) Code of Conduct is a professional code of conduct that all members of the BCS must adhere to. The set of guidelines is in place to ensure that all members conduct themselves in a professional manner and to ensure that all work conducted is ethical and legal. \cite{BSC2022codeofconduct} The BCS Code of Conduct is split into four main sections, which are as follows:

\subsubsection{Public Interest}

The aim of this project is to find the best-performing algorithm with the purpose of creating more benchmarks for the field of multi-objective RL. The research conducted in this project is in the public interest as it aims to further the field of multi-objective RL, which has the potential to be used in a wide range of applications. The four points of the BCS Code of Conduct that are relevant to this project are as follows:

\begin{itemize}
    \item \textbf{Regard for public health, privacy, security and wellbeing} has been taken into account throughout the research conducted. The purpose of the research conducted was to further the field of multi-objective RL which has many potential beneficial applications to society. Moreover, the research conducted was not harmful to any individual or group of individuals.
    
    \item \textbf{Regard for the legitimate rights of third parties} has been taken into account throughout the research conducted. Any third-party work or software used has been explicitly referenced and credited. Lastly, the Pokémon Red ROM was used for educational purposes and not for commercial gain.
    
    \item \textbf{Conduct of professional activities without discrimination} has been taken into account throughout the research conducted. Individuals involved in the research have been treated fairly and equally, and discrimination has not been tolerated.

    \item \textbf{Promote equal access to the benefits of IT} has been taken into account throughout the research conducted. Details of the work conducted have been made public to ensure that the work conducted is transparent and can be reviewed by the public.
\end{itemize}

\subsubsection{Professional Competence and Integrity}

Throughout the research conducted, the researcher has ensured that any claim of work conducted is accurate and truthful. Moreover, the researcher has ensured that the work conducted is of a high standard and that the work conducted is of a professional standard. Furthermore, respect for others' opinions and work has been taken into account throughout the research conducted. 

\subsubsection{Duty to Relevant Authority}

Throughout the research conducted, the researcher has ensured that all professional responsibility has been taken into account. Moreover, the research conducted is not for personal or commercial gain and is for research purposes only. 