\documentclass[11pt]{surrey_disso_style}

\raggedbottom

\usepackage{bookmark}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\usepackage[style=ieee,sorting=none,backend=biber, style=numeric]{biblatex}
\addbibresource{report.bib}

\title{Reinforcement Learning in Pokémon Red to Explore Multi-Objective Environments}
\author{Liam O'Driscoll}
\urn{6640106}
\date{\today}
\degree{Bachelor of Science in Computer Science with a year in placement}
\supervisor{Dr. Sotiris Moschoyiannis}

\begin{document}
\maketitle


\noindent\LARGE\textbf{Abstract}
\normalsize
\vspace{0.5cm}
\newline
This project explores the use of reinforcement learning in Pokémon Red to explore multi-objective environments. The project aims to create a reinforcement learning agent that can learn to play Pokémon Red. The project will use the OpenAI Gymnasium framework to create a custom environment for the agent to interact with. The agent will be trained using the Proximal Policy Optimization (PPO), Deep Q-Network (DQN) and Quantile Regression Deep Q-Network (QRDQN) and compared to determine which algorithm is most effective at multi-objective reinforcement learning. 

The results found that the algorithm PPO performed the best out of the three algorithms followed by QRDQN and DQN last. PPO was able to complete the objective of recieving the first gym badge the earliest and most consistently while maintaining a low consistent low loss level. This demonstrates that policy-based algorithms are far better at solving solving complex multi-objective environments with a large search space compared to value-based algorithms. However, due to the lack of benchmarks for multi-objective reinforcement learning in the field, it is difficult to find further research that agree with this conclusion.

\noindent\LARGE\textbf{Acknowledgements}
\normalsize
\vspace{0.5cm}
\newline
I would like to thank my supervisor, Dr. Sotiris Moschoyiannis, for his guidance and support throughout the project. I would also like to thank my family and friends for their support and encouragement throughout my university experience. Lastly, I would like to thank the University of Surrey for providing me with the opportunity to complete my degree and persue my interest in reinforcement learning.

\newpage

\tableofcontents

\include{1_Introduction}
\include{2_Literature_Review}
\include{3_Research_Problem}
\include{4_Project_Design}
\include{5_Requirements}
\include{6_Implementation_Design}
\include{7_Evaluation}
\include{8_Conclusion}
\include{9_Ethical_Consideration}
\include{10_References}

\end{document}
